Two subtle problems with over-representation analysis

### 2024-06-06 manuscript/manuscript.Rmd
ms

* Reproducibility

* Enrichment analysis

@@

* Docker

@@

Functional enrichment analysis involves the summarisation of omics data to reflect
biological changes, and has become one of the most popular methods in bioinformatics.
Enrichment analysis software packages have been collectively cited 131,332 times as of
late 2019 [@Xie2021-ch].
The algorithms used by these tools can largely be divided into two groups, functional
class scoring (FCS) and over-representation analysis (ORA).
In FCS, all detected genes are ranked by their degree of differential expression followed
by some statistical test for enrichment of gene categories at either extreme of the
ranked list.
ORA involdes the selection of genes of interest, followed by a test to ascertain whether
certain functional categories are over-represented in the selected genes.

Of these two methods, FCS is reported to be the most accurate, while ORA seems to be the
most popular.

According to a study that used simulated differential expression data, FCS accuracy was
superior to ORA over a range of experimental designs [@Kaspi2020-rn].

A survey of articles published in 2019 showed that seven of the top eight most popular
methods were based on ORA.

>>>>

@@

* Research integrity

@@

Functional enrichment analysis involves the summarisation of omics data to reflect
biological differences, and has become one of the most popular methods in bioinformatics.
Enrichment analysis software packages have been collectively cited 131,332 times as of
late 2019 [@Xie2021-ch].
The algorithms used by these tools can largely be divided into two groups, functional
class scoring (FCS) and over-representation analysis (ORA).
In FCS, all detected genes are ranked by their degree of differential expression followed
by a statistical test for enrichment of gene categories at either extreme of the
ranked list.
ORA involdes the selection of genes of interest, followed by a test to ascertain whether
certain functional categories are over-represented in the selected genes.

According to a study that used simulated differential expression data, FCS accuracy was
superior to ORA over a range of experimental designs [@Kaspi2020-rn].
Despite this performance difference, a survey of articles published in 2019 showed that
seven of the top eight most popular tools were based on ORA [@Wijesooriya2022-fw].
The discrepancy between performance and popularity likely arose due to the relative ease of
conducting ORA as compared to FCS.
A minimalist ORA might involve pasting a single list of genes into a text box on a website,
where the results appear nearly instantaneously.
On the other hand, FCS typically involves downloading specific softwares and dealing with a
dataset that represents every detected gene, typically ~20,000 rows and ensuring that the
format from upstream tools can be shoehorned to meet the requirements of FCS packages.

Since it's initial development in 1999 [@Tavazoie1999-bu], ORA has proliferated widely,
becoming part of many websites and software packages.
The most popular ORA website is DAVID [@Dennis2003-ar;@Huang2009-gi;@Sherman2022-jn],
followed by Ingenuity Pathway Analysis, a commercial software package provided by QIAGEN Inc.
ORA has been implemented into popular R/Bioconductor packages including `ClusterProfiler`
[@Yu2012-jj],`limma`[@Ritchie2015-oz] and `GOseq`[@Young2010-iw].

Despite the popularity of the technique, there are concerns that
